Simple models of deliberated preference

### 2017-11-06 article.tex
¬YM
Started improving basic defs

\RequirePackage[l2tabu, orthodox]{nag}
\RequirePackage{silence}\WarningFilter{fmtcount}{\ordinal already defined use \FCordinal instead}
\documentclass[version=last, pagesize, twoside=semi, DIV=calc, bibliography=totoc, 12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\input{preamble/draw}
\input{preamble/acronyms}

\begin{document}
\title{Simple models of deliberated preference}
\author{Olivier Cailloux \and Yves Meinard}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

@@

\appendix
\section{Todo}
\begin{itemize}
	\item If $i$ does not consider $\ar$ as supporting $\prop$, it also works: if $\prop$ is not weakly acceptable by default, then any $\ar'$ is considered by $i$ as a better argument than $\ar$ in favor of certain $¬\prop$, and so on. In fact, whether $\emptyset \ibeatse[\prop] \emptyset$ determines whether $\prop$ is weakly supported by default.
	\item I should define $\ar' (□\ibeatse[\prop]) \ar$ as an observable: “Assuming $\ar'$ would survive, do you consider $\ar'$ as leading to certainty of $¬\prop$, even when considering $\ar$?”. It distinguishes our knowledge and the truth: $\ar' (□\ibeatse[\prop]) \ar ⇒ \ar' \ibeatse[\prop] \ar$, thus, implies $¬ (\ar' \nibeatsst[\prop] \ar)$. But out of $¬ (\ar' (□\ibeatse[\prop]) \ar)$, nothing.
	\item Partition (objectively) $\allargs$ (or $\allargs × \alltopic$) into arguments in favor of $\prop, \text{sure}$, $¬\prop, \text{sure}$, and similarly for possible. Use only one rel $\ibeatse$, defined on contradictory arguments only, instead of $\ibeatse[\prop, \text{sure}]$ and others. Define $\ar' \ibeatse[\prop] \ar$ equals no when $¬(\ar' \ileadsto ¬\prop, \text{sure})$, equals $\ibeatse$ for adequate arguments, and equals yes when $¬(\ar \ileadsto \prop, \text{possible})$ and $\ar' \ileadsto ¬\prop, \text{sure}$, with probably some complications needed for the argument $\emptyset$ (and related default attitude towards $\prop$).
\end{itemize}

>>>>

\RequirePackage[l2tabu, orthodox]{nag}
\RequirePackage{silence}\WarningFilter{fmtcount}{\ordinal already defined use \FCordinal instead}
\documentclass[version=last, pagesize, twoside=semi, DIV=calc, bibliography=totoc, 12pt, a4paper, french, english]{scrartcl}
\input{preamble/packages}
\input{preamble/math_basics}
\input{preamble/math_mine}
\input{preamble/redac}
\input{preamble/draw}
\input{preamble/acronyms}
\addtokomafont{labelinglabel}{\sffamily\bfseries}

\begin{document}
\title{Simple models of deliberated preference}
\author{Olivier Cailloux}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

\subsection{Possible outcomes}
We know $\prop \in \isure ⇒ ¬\prop \notin \iposs$ and $\prop \in \isure ⇒ \prop \in \iposs$.

Here are the remaining possibilities, considering the four pairs $(\prop, \iposs)$, $(\prop, \isure)$, $(¬\prop, \iposs)$, $(¬\prop, \isure)$. (A pair $(\prop, T)$ not mentioned means $\prop \notin T$.)
\begin{labeling}{Unstability with $¬\prop$ possible}
	\item[All poss] $\prop \in \iposs$, $¬\prop \in \iposs$
	\item[$\prop$ sure] $\prop \in \iposs$, $\prop \in \isure$
	\item[$¬\prop$ sure] (symmetric)
	\item[Unstability with $\prop$ possible] $\prop \in \iposs$ (oddity: $¬\prop \notin \iposs$ should imply $\prop \in \isure$)
	\item[Unstability with $¬\prop$ possible] (symmetric)
	\item[Strong unstability] (both oddities above)
\end{labeling}

@@

\appendix
\section{Better definitions}
Define $f_\prop$ from any unordered pair of arguments $(\ar, \ar')$ into $\{$ $\prop$ sure, $¬\prop$ sure, both possible, unstable$\}$. We present both arguments to $i$ and ask which proposition $i$ considers valid in her current state of mind, thus using both arguments and possibly other arguments she has in mind, where the first three are when $i$ always declares the same answer and the last one is used if $i$ sometimes wobble (because she has learnt new things in between two questioning, or just because of time passing, or for any known or unknown reason).

We can also use $\emptyset$ as one or both of the arguments.

We write $(\ar, \ar') \ileadsto \prop$ (respectively $¬\prop$, $b$, $\emptyset$) for $f_\prop(\ar, \ar') = \prop \text{ sure}$ (respectively $¬\prop$ sure, both possible, unstable).

To encode this data efficiently, we can define two binary relations $\prop^+$ and $\prop^−$: $(\ar, \ar') \in \prop^+$ ⇔ $(\ar, \ar')$ always supports $\prop$ sure or always supports both ⇔ $(\ar, \ar') \ileadsto \prop ∨ (\ar, \ar') \ileadsto b$; and $(\ar, \ar') \in \prop^− ⇔ (\ar, \ar') \ileadsto ¬\prop ∨ (\ar, \ar') \ileadsto b$.

This is indeed equivalent:
\begin{itemize}
	\item $(\ar, \ar') \ileadsto \prop ⇔ (\ar, \ar') \in \prop^+ ∧ (\ar, \ar') \notin \prop^−$
	\item $(\ar, \ar') \ileadsto ¬\prop ⇔ (\ar, \ar') \notin \prop^+ ∧ (\ar, \ar') \in \prop^−$
	\item $(\ar, \ar') \ileadsto b ⇔ (\ar, \ar') \in \prop^+ ∧ (\ar, \ar') \in \prop^−$
	\item $(\ar, \ar') \ileadsto \emptyset ⇔ (\ar, \ar') \notin \prop^+ ∧ (\ar, \ar') \notin \prop^−$
\end{itemize}

We can now define $\ipossp, \ipossm \subseteq \alltopic$:
\begin{itemize}
	\item $\prop \in \ipossp ⇔ \exists \ar \suchthat \forall \ar': [(\ar, \ar') \ileadsto \prop] ∨ [(\ar, \ar') \ileadsto b]$. 
	\item $\prop \in \ipossm ⇔ \exists \ar \suchthat \forall \ar': [(\ar, \ar') \ileadsto ¬\prop] ∨ [(\ar, \ar') \ileadsto b]$. 
	\item $\prop$ strongly possible ⇔ $[\exists \ar \suchthat \forall \ar': (\ar, \ar') \ileadsto \prop] ∨ [\exists \ar \suchthat \forall \ar': (\ar, \ar') \ileadsto b]$. 
\end{itemize}

Here is a case where $\prop$ is possible but not strongly possible: $\exists (\ar_1, \ar) \ileadsto \prop$, $\exists (\ar_2, \ar) \ileadsto b$, $\nexists (\ar', \ar) \ileadsto ¬\prop$, $\nexists (\ar', \ar) \ileadsto \emptyset$. However this case is odd as just after showing $\ar_2$ to $i$ we could ask again about $\ar_1$ … (TODO)

Now do we have $\prop \in \isure ⇔ \exists \ar \suchthat \forall \ar': (\ar, \ar') \ileadsto \prop$?
\section{Todo}
\begin{itemize}
	\item If $i$ does not consider $\ar$ as supporting $\prop$, it also works: if $\prop$ is not weakly acceptable by default, then any $\ar'$ is considered by $i$ as a better argument than $\ar$ in favor of certain $¬\prop$, and so on. In fact, whether $\emptyset \ibeatse[\prop] \emptyset$ determines whether $\prop$ is weakly supported by default.
	\item I should define $\ar' (□\ibeatse[\prop]) \ar$ as an observable: “Assuming $\ar'$ would survive, do you consider $\ar'$ as leading to certainty of $¬\prop$, even when considering $\ar$?”. It distinguishes our knowledge and the truth: $\ar' (□\ibeatse[\prop]) \ar ⇒ \ar' \ibeatse[\prop] \ar$, thus, implies $¬ (\ar' \nibeatsst[\prop] \ar)$. But out of $¬ (\ar' (□\ibeatse[\prop]) \ar)$, nothing.
	\item Partition (objectively) $\allargs$ (or $\allargs × \alltopic$) into arguments in favor of $\prop, \text{sure}$, $¬\prop, \text{sure}$, and similarly for possible. Use only one rel $\ibeatse$, defined on contradictory arguments only, instead of $\ibeatse[\prop, \text{sure}]$ and others. Define $\ar' \ibeatse[\prop] \ar$ equals no when $¬(\ar' \ileadsto ¬\prop, \text{sure})$, equals $\ibeatse$ for adequate arguments, and equals yes when $¬(\ar \ileadsto \prop, \text{possible})$ and $\ar' \ileadsto ¬\prop, \text{sure}$, with probably some complications needed for the argument $\emptyset$ (and related default attitude towards $\prop$).
\end{itemize}
